Technology-based aids for people affected by Autism Spectrum Disorder (ASD)

### 2017-11-16 paper.tex
Added frist part on hand eye coordination

The classical way to conduct a therapy session, is a face-to-face meeting between the subject and the therapist.
Usually, is conducted in four phases: the first one is called \textit{Instruction}, in which the therapist explain what is the skill that the subject is going to learn; the second one is the \textit{Modeling} phase, in which the therapist show the patient what he has to do; there is then the \textit{Rehearsal} phase, in which the patient tries to imitate the therapist and finally a \textit{Feedback} phase in which the therapist draw its own conclusions about the test.
The problem of this kind of therapy are the completely subjective evaluation procedure carried out by the therapist and that the subject's attention is not immediate.

@@

>>>>

The classical way to conduct a therapy session, is a face-to-face meeting between the subject and the therapist.
Usually, it is conducted in four phases: the first one is called \textit{Instruction}, in which the therapist explains what is the skill that the subject is going to learn; the second one is the \textit{Modeling} phase, in which the therapist shows the patient what he has to do; there is then the \textit{Rehearsal} phase, in which the patient tries to imitate the therapist and finally a \textit{Feedback} phase in which the therapist draws its own conclusions about the test.
The problem of this kind of therapy are the completely subjective evaluation procedure carried out by the therapist and that the subject's attention is not immediate.

@@

\section{Diagnostic techniques}
\label{sec:diagnosis}

\subsection{Hand-eye coordination}
\label{sec:hand_eye}
A different aspect to take into consideration when facing subjects with ASD, is the motor disorder related to the syndrome. In \cite{Casellato2017} a new approach to face the issue is presented, where the proposed protocol may be used for ad hoc subject training, helping children with ASD to enhance motor abilities related to hand-eye coordination.

Moreover, although normally it is not considered as a diagnostic criteria, motor impairment associated with ASD may have a significant impact on daily life tasks and, since up to 80\% of children with autism show motor anomalies highly correlated with the severity of social deficiency, hand-eye coordination could be considered as a new early diagnostic strategy for ASD syndrome \cite{Casellato2017}. In fact, it has been suggested that sensory integration with bad timing brings to poor motor performance in children with ASD: the task of integrating multisensory input influences the forming of coherent perception, action coordination and planning. Nowadays, the normal procedure to assess this motor coordination deficiency is the clinical observation and self-report questionnaires, so a new non-invasive objective method to map the development of ASD such hand-eye motor performance, could be a powerful diagnostic tool.

\subsubsection{Material and methods}
In \cite{Casellato2017} the experimental set-up involved sixteen children aged between 8 and 10 years old and was composed by a haptic robotic manipulandum (Phantom Omni, Sensable-Geomagic) with a 3D printed pen-like handle, a control pc with a suited control algorithm running and a table-mounted screen Eye tracker T60 (Tobii) of $1280\times1024$ pixels. The manipulandum's sampling frequency was of 1 kHz, while the Eye Tracker's sampling frequency was of 60 Hz.

Prior to the beginning of the trial, a brief amount of time was given to the children in order to familiarize with the system.

The actual trial consisted in two main tasks: the first task was about following a diagonal straight path moving an end-effector from a start square to an end square without exiting the path boundaries; in the second task a target moved on the screen following a curvilinear predefined path representing a digitalized version of the visuo-motor precision sub-task of the NEPSY-II protocol.

During the first task, a primary amount of attempts was done without applying external force disturbance to the pen-like handle; the attempts that followed, instead, were perturbed by an external force, while the final ones were not perturbed again. Results from this task show that:
\begin{itemize}
\item the task was fulfilled pretty well in the first unperturbed part;
\item during the first perturbed attempts the performance of the subjects were significantly low and attest a slow adaptation to the change;
\item after some perturbed attempts the subjects eventually manage to compensate the force almost properly, with a progressive reduction in the product $pathL \dot T$ (where $pathL$ is the length of the path while $T$ is the time taken to travel through it);
\item when the attempts go back to be not-perturbated there is an initial overcompensation, which eventually is re-calibrated with even better performance than the ones obtained initially.
\end{itemize}

Figures \ref{task1} and \ref{task1_pT} show the results just described.

\begin{figure}[h]
\centering
\begin{minipage}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{task1.png}
\caption{Task 1 path following}
\label{fig:task1}
\end{minipage}
\begin{minipage}{0.48\textwidth}
\centering
\includegraphics[width=\textwidth]{task1_pT.png}
\caption{Task 1 pathLxT vs n of trials}
\label{fig:task1_pT}
\end{minipage}
\end{figure}
